---
title: "Transmission Type vs Fuel Efficiency"
author: "by Rajesh Sankar"
output:
  pdf_document:
    fig_height: 7
    fig_width: 7
---

## Executive Summary  
In this article, we will look at the effect of automatic vs manual transmission on fuel efficiency. To examine, we use a dataset for 32 automobiles (all 1973 - 1974 models) which includes a wide range of data comprising of mpg, number of cylinders, horse power etc. 

## Exploratory Data Analysis  

### Data Set
First, we look at the data set. This data set was extracted from the 1974 edition of Motor Trend magazine. It consists of 32 cars on following 11 different variables:

- Miles per US gallon
- Number of cylinders
- Displacement (cubic inches)
- Horsepower
- Rear axle ratio
- Weight (lb / 1000)
- 1/4 mile time
- V/S
- Transmission (automatic, manual)
- Number of Gears
- Number of carburetors

The exploratory analysis is described in **Appendix** section for the plots. The box plot shows hat on average there is a difference between fuel efficiency depending on the transmission type. The manual transmission yields higher MPG than the automatic transmission. The Diagnostic plots show Toyota Corolla and Fiat 128 have a very high fuel efficiency, while Chrysler Imperial has a low fuel efficiency. Toyota Corona has a medium fuel efficiency. The pair graph shows some higher correlations between Weight, Displacement, number of Cylinders and Horsepower variables.

## Regression Model
Stepwise model selection using backwards elimination was performed to determine the variables for the best model.
```{r}
full.model <- lm(mpg ~ ., data = mtcars)
best.model <- step(full.model, direction = "backward")
```

### Summary
```{r}
summary(best.model)
```

### Result
Thus, this regression model  shows the best model includes cyl6, cyl8, hp, wt, and amManual variables. The adjusted R-squared indicates that about 84% of the variance is explained by the final model. The MPG decreases with higher number of cylinders, horsepower and weight. MPG increases with manual transmission.

## Appendix: Graphs

### Boxplot: Transmission vs MPG
```{r, fig.width=5, fig.height=5}
boxplot(mpg ~ am, data=mtcars, xlab="Transmission", ylab="MPG", col=terrain.colors(3))
title(main = "Transmission vs MPG")
legend("topleft", inset=.03, c("automatic", "manual"), fill=terrain.colors(3),horiz=TRUE)
```

### Pair Graph
```{r}
pairs(mtcars, panel=panel.smooth, main="Pair Graph")
``` 

### Residual Plots
```{r}
par(mfrow = c(2, 2))
plot(lm(mpg ~ wt + qsec + am + wt:am, data=mtcars))
```

